---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

```{r setup, echo=FALSE}
library(ribbon)
library(knitr)
library(ggplot2)
library(sf)
library(tidyverse)
library(grid)
library(viridis)
```
  
### Overview

ribbon provides ggplot2 extension to produce glyph plots with `geom_glyph_ribbon()` and 
`geom_segment_glyph()`.

These functions create a ribbon geometry designed to display glyphs
based on the combination of `x_major` and `y_major`. For each `x_minor`
value, `geom_glyph_ribbon()` displays a y interval defined by
`ymin_minor` and `ymax_minor`. 

### Installation

```{r, eval=FALSE}
# Download the development version from GitHub:
devtools::install_github("maliny12/ribbon")
```

### Options

Options allow you to modify how functions behave by altering parameters to meet specific needs. All of which are global options that affect all glyphs.

|Option          | Default      | Description
|--------------- | ------------ | ------------------------------------------------
| `x_scale`         | `"identity"`         | This function scales each set of minor values within a grid cell for the x-dimension.
|`y_scale`      | `"identity"`          | This function scales each set of minor values within a grid cell for the y-dimension.
|`width`         | `0.1`          |  Specifies the width of the glyph.
|`height`     | `0.1`     | Specifies the height of the glyph.
|`global_rescale`      | `TRUE`        | Determines whether rescaling is performed globally across all glyphs or individually for each glyph.

### Aesthetics

Aesthetics are parameters that you can map to your data using geom_glyph_ribbon(mapping = aes(...)).

`ribbon` provides the same aesthetics for geom_glyph_ribbon and geom_segment_glyph() as are available in geom_ribbon() or geom_segment(), but it also introduces additional unique options.

|Option          | Default      | Description
|--------------- | ------------ | ------------------------------------------------
| `colour`         | `"grey50"`         | Color for line segments and borders.
|`linewidth`      | `0.5`          | Width of the line for borders.
|`linetype`         | `1`          |  Style of the line for borders.
|`fill`     | `"grey40"`     | Color of the interior area of the geometries.
|`alpha`      | `1`        | Transparency level of the glyphs. 


### Examples
#### Comparison between global rescale and local rescale

(......)

```{r}

p1 <- aus_temp |>
   ggplot(aes(x_major = long, y_major = lat,
              x_minor = month, ymin_minor = tmin, ymax_minor = tmax)) +
  geom_sf(data = ozmaps::abs_ste,
          fill = "grey95", color = "white",
          inherit.aes = FALSE)  +
   add_glyph_boxes() +
   add_ref_lines() +
  geom_glyph_ribbon(global_rescale = FALSE) +
  theme_bw() +
  coord_sf(xlim = c(110,155)) +
  labs(title = "Local Rescale",
       x = "Longitude", y = "Latitude") 

p2 <- aus_temp |>
   ggplot(aes(x_major = long, y_major = lat,
              x_minor = month, ymin_minor = tmin, ymax_minor = tmax)) +
  geom_sf(data = ozmaps::abs_ste,
          fill = "grey95", color = "white",
          inherit.aes = FALSE)  +
   add_glyph_boxes() +
   add_ref_lines() +
  geom_glyph_ribbon(global_rescale = TRUE) +
  theme_bw() +
  coord_sf(xlim = c(110,155)) +
  labs(title = "Global Rescale",
       x = "Longitude", y = "Latitude") 

gridExtra::grid.arrange(p1, p2, ncol = 2) 

```

#### Mapping precipitation across Australia with glyphs

(......)

```{r}

aus_temp |>
   group_by(id) |>
   mutate(prcp = mean(prcp, na.rm = TRUE)) |>
   ggplot(aes(x_major = long, y_major = lat, x_minor = month,
              ymin_minor = tmin, ymax_minor = tmax,
              fill = prcp, color = prcp)) +
  geom_sf(data = ozmaps::abs_ste,
           fill = "grey95", color = "white",
           inherit.aes = FALSE)  +
   add_glyph_boxes() +
   add_ref_lines() +
   geom_glyph_ribbon() +
   coord_sf(xlim = c(110,155)) +
  theme_bw() +
  labs(x = "Longitude", y = "Latitude")

```

#### Comparative visualization by factors across years

(......)

```{r}

historical_temp |>
  filter(id %in% c("ASN00026021", "ASN00085291", "ASN00084143")) |>
   ggplot(aes(color = factor(year), fill = factor(year), group = interaction(year,id),
              x_major = long, y_major = lat, x_minor = month, alpha = 0.5,
              ymin_minor = tmin, ymax_minor = tmax)) +
  geom_sf(data = ozmaps::abs_ste |> filter(NAME == "Victoria"),
           fill = "grey95", color = "white",
           inherit.aes = FALSE)  +
   add_glyph_boxes() +
   add_ref_lines() +
   geom_glyph_ribbon() +
   coord_sf(xlim = c(140,150)) +
  labs(x = "Longitude", y = "Latitude") +
  theme_bw()

```

#### Integrating legends in map visualizations

(......)

```{r, asepx}
aus_temp |>
   ggplot(aes(x_major = long, y_major = lat,
              x_minor = month, ymin_minor = tmin, ymax_minor = tmax)) +
  geom_sf(data = ozmaps::abs_ste,
          fill = "grey95", color = "white",
          inherit.aes = FALSE)  +
  add_glyph_boxes() +
  add_ref_lines() +
  geom_glyph_ribbon() + 
  add_ribbon_legend() + 
  theme_glyph()
```

